Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-11-14 review/response.md

4.2 General Comments
--------------------

Some areas where I would like to see more detail:

1. It would have been interesting to understand the temporal characteristics for
a week pedestrian footfall. How to consider this situation in your study? Why do
you choose time range from 12:30 to 13:00 hrs ?

> Thank you. Our related work (cited below) examines daily, weekly and seasonal
> variation, and wil be the subject of future papers.

> Lugomer, K., & Longley, P. (2018). Towards a Comprehensive Temporal 
> Classification of Footfall Patterns in the Cities of Great Britain (Short
> Paper). In LIPIcs-Leibniz International Proceedings in Informatics (Vol. 114).
> Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik.

> There was no reason for
> selecting the time in the pilot study. the location was chosen for it it known
> to have high number of randomised probe requests and has one of the highest
> footfall in London. For the main study the counting was done such that each
> location has atleast 2/3 counts in the same time interval.

@@

4.3 Corrections and Clarifications
----------------------------------

More specific comments can be found in the following.

1. Minor formatting issues and typos needs to be fixed.
Figure 2, no value and units in the x-axis and y-axis.
Table 1, and Figure 4 are in wrong location.

> Thank you. This has been corrected in the revised version of the manuscript.
> Figure 2 is just a schematic showing the idea of cut off and clustering hence
> there are now values on the axes. We have included the units in the description

### 2018-11-14 review/response.md

5. P7 Section 3.3, please be more specific about the calibration process and 
the external source of information to be used here.

> This section has been expanded to include more details. The calibration process
> is adjusting the future counts based on a sample manual counts done at the
> location. The external source of information used is the manual counts which is
> collected outside the sensor

6. P7 l45, the analysis in this research is based on one single sensor for each 
location; however, in reality, it is very common to estimate travel flows based 
on multiple sensors (e.g., nearby stores in a shopping mall). How would your 
method address duplication in multiple sensors?

> This is a valid concern. Which we tried to clarify in the revised manuscript.
> Our motivation is not to attempt to track the movements of individuals,
> although we do attempt to accommodate measurement variations attributable to 
> micro-site location of the sensors. In related work we have carried out 
> controlled comparison of sensor measures obtained from adjacent store
> locations, and examined correspondence with retail unit function. This enables
> identification of the differences in dwell time between, for example,
> fast food outlets and clothes shops. This full analysis requires
> addition of store till data and the analysis is detailed. For this
> reason we include only a short reference to this extension of our
> analysis here. We also aim this method to count footfall accurately at 
> locations rather than the flow of footfall between those locations. But as 
> demonstarted in

> Lee J, Nemati S, Silva I, Edwards BA, Butler JP, Malhotra A. Transfer entropy 
> estimation and directional coupling change detection in biomedical time series.
> Biomed Eng Online. 2012;11:19. Published 2012 Apr 13.
> doi:10.1186/1475-925X-11-19

> and

> Vicente R, Wibral M, Lindner M, Pipa G. Transfer entropy--a model-free 
> measure of effective connectivity for the neurosciences.
> J Comput Neurosci. 2010;30(1):45-67.

> we can infer the relation or flow between the locations using other
> probabilistic approaches.

@@

> Thank you. Our related work (cited below) examines daily, weekly and seasonal
> variation, and wil be the subject of future papers.

> Lugomer, K., & Longley, P. (2018). Towards a Comprehensive Temporal 
> Classification of Footfall Patterns in the Cities of Great Britain (Short
> Paper). In LIPIcs-Leibniz International Proceedings in Informatics (Vol. 114).
> Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik.

> There was no reason for
> selecting the time in the pilot study. the location was chosen for it it known
> to have high number of randomised probe requests and has one of the highest
> footfall in London. For the main study the counting was done such that each
> location has atleast 2/3 counts in the same time interval.

2. Will you give more detailed information about the data and the parameters for 
various clustering algorithms? Including the number of records, the number of 
clusters. How about the performance for different clustering methods 
in your study？

### 2018-11-18 paper/sections/pilot_study.tex
Thu 15 Nov 15:14:58 GMT 2018
fixes

The next challenge was to identify probe requests which are generated by the same device irrespective of MAC randomisation process.
We use the algorithm defined in Chapter \ref{methodology} and assign a unique identifier or signature to each probe request independent of the MAC address. 
Since there is neither prior research nor documentation on a universal behaviour of phones in randomising their MAC addresses, we use the stationary device - the one used for manual counting as a reference and find out the suitable time threshold, $\alpha$ and threshold for sequence numbers, $\beta$ to be 16s and 60 respectively via trial and error.
This process is done on top the filtering done based on signal strength and only for the probe requests with randomised MAC addresses.
Figure \ref{pilot_results} shows the results of this clustering process on a small set of randomised probe requests.
The probe request with different randomised MAC address is shown by the colored points and the line joining them shows the ones belonging to the same cluster hence expected to be generated by the same device.
We finally aggregate the probe requests as before but with the device signature rather than just MAC addresses this results in a footfall count with a MAPE of -18\%. 
A comparison of minute by minute counts resulting from different filtering processes along with the ground truth is shown in Figure \ref{pilot_results} showing the promising effectiveness of the methods.

>>>>

The next challenge was to identify probe requests which are generated by the same device irrespective of MAC randomisation process.
We use the algorithm defined in Chapter \ref{methodology} and assign a unique identifier or signature to each probe request independent of the MAC address. 
Since there is neither prior research nor documentation on a universal behaviour of phones in randomising their MAC addresses, we use the stationary device - the one used for manual counting as a reference and find out the suitable time threshold, $\alpha$ and threshold for sequence numbers, $\beta$ to be 16 seconds and 60 respectively via trial and error.
This process is done on top the filtering done based on signal strength and only for the probe requests with randomised MAC addresses.
Figure \ref{pilot_results} shows the results of this clustering process on a small set of randomised probe requests.
The probe request with different randomised MAC address is shown by the colored points and the line joining them shows the ones belonging to the same cluster hence expected to be generated by the same device.
We finally aggregate the probe requests as before but with the device signature rather than just MAC addresses this results in a footfall count with a MAPE of -18\%. 
A comparison of minute by minute counts resulting from different filtering processes along with the ground truth is shown in Figure \ref{pilot_results} showing the promising effectiveness of the methods.
